Manuscript Figures for Genome Biology (simulation subset)

### 2018-09-06 manuscript_figures/manuscript_figures_sims.Rmd
Update and clean simulation main and suppl plots

```{r}
genplot <- function(tab, cov, xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
    met <- match.arg(met)
    cov <- rlang::enquo(cov)
    gp <- dplyr::filter(tab, key == met) %>%
        ggplot(aes(x = !!cov, y = mean.info, color = Method, group = Method)) +
        geom_line(aes(linetype = Method), alpha = 0.85) +
        geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = ebw, alpha=0.5) +
        scale_linetype_manual(values = lty) + 
        scale_color_manual(values = col) +
        scale_y_continuous(met, labels = scales::percent) +
        expand_limits(y = 0) +
        ggtitle(paste("Mean", met)) +
        theme_classic() +
        theme(axis.title = element_text(face = "bold", size = 10),
              plot.title = element_text(face = "bold", size = 14))
    if (met == "FDR") {
        if (rlang::quo_name(cov) == "alpha") {
            gp <- gp + geom_abline(lty = 2, color = "blue", alpha = 1/2) 
        } else {
            gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
        }
    }
    gp
}
```

@@

```{r, fig.width = 10, fig.height = 12}
## Figure 3A,B: FDR/TPR of methods across varying number of tests at alpha = 0.05
res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
p3a <- genplot(res_ntests, cov = as.factor(ntests), met = "FDR", ebw = 0.15) +
    xlab("Number of Tests")
p3b <- genplot(res_ntests, cov = as.factor(ntests), met = "TPR", ebw = 0.15) +
    xlab("Number of Tests")

## Figure 3C,D: FDR/TPR of methods across varying null proportion at alpha = 0.05
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
p3c <- genplot(res_pi0, cov = pi0, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Proportion of Null Hypotheses",
                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
p3d <- genplot(res_pi0, cov = pi0, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Proportion of Null Hypotheses",
                       breaks = c(5, seq(10, 90, by = 10), 95, 99))

## Figure 3E,F: FDR/TPR of methods across varying informativeness at alpha = 0.05
res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
p3e <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
p3f <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))

Fig3 <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
                p3b + guides(color = FALSE, linetype = FALSE),
                p3c + guides(color = FALSE, linetype = FALSE),
                p3d + guides(color = FALSE, linetype = FALSE),
                p3e + guides(color = FALSE, linetype = FALSE),
                p3f + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:6], ncol = 2)
Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
Fig3
ggsave(file.path(outdir, "Figure3.pdf"), width=10, height=12)
```

# Figure 3+

This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different test statistic distributions. Test statistics were simulated from Normal, t with 11 degrees of freedom, t with 5 degrees of freedom, and Chi-squared with 4 degrees of freedom.

```{r, fig.width = 10, fig.height = 12}
## Figure 3plus A,B: FDR/TPR of methods across varying test statistics distribution
res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
p3pa <- genplot(res_noise, cov = alpha, met = "FDR") +
    facet_grid(. ~ dist, labeller = label_parsed)
p3pb <- genplot(res_noise, cov = alpha, met = "TPR") +
    facet_grid(. ~ dist, labeller = label_parsed)

Fig3p <- plot_grid(p3pa + guides(color = FALSE, linetype = FALSE),
                   p3pb + guides(color = FALSE, linetype = FALSE),
                   labels = LETTERS[1:2], ncol = 1)
Fig3p <- plot_grid(Fig3p, get_legend(p3pa), rel_widths = c(7.5, 1))
Fig3p
ggsave(file.path(outdir, "Figure3plus.pdf"), width=14, height=8)
```

>>>>

```{r genplot-function}
genplot <- function(tab, cov, type = c("info", "diff"), xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
    met <- match.arg(met)
    cov <- rlang::enquo(cov)
    type <- match.arg(type)
    if (type == "info") {
        ymean <- rlang::quo(mean.info)
        yse <- rlang::quo(se.info)
    } else if (type == "diff") {
        ymean <- rlang::quo(mean.diff)
        yse <- rlang::quo(se.diff)
    }        
    gp <- dplyr::filter(tab, key == met) %>%
        ggplot(aes(x = !!cov, y = !!ymean, color = Method, group = Method)) +
        geom_line(aes(linetype = Method), alpha = 0.85) +
        geom_errorbar(aes(ymin = !!ymean - !!yse, ymax = !!ymean + !!yse), width = ebw, alpha=0.5) +
        scale_linetype_manual(values = lty) + 
        scale_color_manual(values = col) +
        scale_y_continuous(met, labels = scales::percent) +
        expand_limits(y = 0) +
        theme_classic() +
        theme(axis.title = element_text(face = "bold", size = 10),
              plot.title = element_text(face = "bold", size = 14))
    if (type == "diff") {
        gp <- gp + geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) 
    } else if (met == "FDR") {
        if (rlang::quo_name(cov) == "alpha") {
            gp <- gp + geom_abline(lty = 2, color = "blue", alpha = 1/2) 
        } else {
            gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
        }
    }
    gp
}
```

@@

```{r Figure3, fig.width = 10, fig.height = 12}
## Figure 3A,B: FDR/TPR of methods across varying informativeness at alpha = 0.05
res_info <- dplyr::filter(resmet, setting == "varyinginfo-smooth", alpha == 0.05)
p3a <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
p3b <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))

## add title to 3A,B
p3ab <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
                  p3b + guides(color = FALSE, linetype = FALSE),
                  labels = LETTERS[1:2], ncol = 2)
p3abTitle <- ggdraw() +
    draw_label("Simulation performance across varying informativeness",
               fontface = 'bold')
p3ab <- plot_grid(p3abTitle, p3ab, ncol = 1, rel_heights = c(0.1, 1))

## Figure 3C,D: FDR/TPR of methods across varying number of tests at alpha = 0.05
res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
p3c <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.05) +
    scale_x_continuous("Number of Tests", trans = "log10",
                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))
p3d <- genplot(res_ntests, cov = ntests, met = "TPR", ebw = 0.05) +
    scale_x_continuous("Number of Tests", trans = "log10",
                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))

## add title to 3C,D
p3cd <- plot_grid(p3c + guides(color = FALSE, linetype = FALSE),
                  p3d + guides(color = FALSE, linetype = FALSE),
                  labels = LETTERS[3:4], ncol = 2)
p3cdTitle <- ggdraw() +
    draw_label("Simulation performance across varying numbers of hypotheses",
               fontface = 'bold')
p3cd <- plot_grid(p3cdTitle, p3cd, ncol = 1, rel_heights = c(0.1, 1))

## Figure 3E,F: FDR/TPR of methods across varying null proportion at alpha = 0.05
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
p3e <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Proportion of Non-Null Hypotheses",
                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))
p3f <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
    scale_x_continuous("Proportion of Non-Null Hypotheses",
                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))

## add title to 3E,F
p3ef <- plot_grid(p3e + guides(color = FALSE, linetype = FALSE),
                  p3f + guides(color = FALSE, linetype = FALSE),
                  labels = LETTERS[5:6], ncol = 2)
p3efTitle <- ggdraw() +
    draw_label("Simulation performance across varying non-null proportion",
               fontface = 'bold')
p3ef <- plot_grid(p3efTitle, p3ef, ncol = 1, rel_heights = c(0.1, 1))

## pull Figure 3 together
Fig3 <- plot_grid(p3ab, p3cd, p3ef, ncol = 1)
Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
Fig3
ggsave(file.path(outdir, "Figure3.pdf"), width=10, height=12)
```

# Supplementary Figure (distributions)

This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different test statistic distributions. Test statistics were simulated from Normal, t with 11 degrees of freedom, t with 5 degrees of freedom, and Chi-squared with 4 degrees of freedom. We plot TPR and FDR across 100 replicates over nominal FDR cutoffs of 0.01 to 0.10. Additionally, we plot the differences in TPR and FDR for methods when using an informative or an uninformative covariate.

```{r SFigureSim1, fig.width = 14, fig.height = 16}
## SFigure Sim1 A,B: FDR/TPR of methods across varying test statistics distribution
res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
ps1a <- genplot(res_noise, cov = alpha, met = "FDR") +
    facet_grid(. ~ dist, labeller = label_parsed)
ps1b <- genplot(res_noise, cov = alpha, met = "TPR") +
    facet_grid(. ~ dist, labeller = label_parsed)

## SFigure Sim1 C,D: FDR/TPR difference of methods
ps1c <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
    facet_grid(. ~ dist, labeller = label_parsed)
ps1d <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
    facet_grid(. ~ dist, labeller = label_parsed)

## create title for SFigure Sim1
ps1Title <- ggdraw() +
    draw_label("Simulation performance across test statistic distributions",
               fontface = 'bold', size = 20)

## pull SFigure Sim1 together
SFigS1 <- plot_grid(ps1a + guides(color = FALSE, linetype = FALSE),
                    ps1b + guides(color = FALSE, linetype = FALSE),
                    ps1c + guides(color = FALSE, linetype = FALSE),
                    ps1d + guides(color = FALSE, linetype = FALSE),
                    labels = LETTERS[1:4], ncol = 1)
SFigS1 <- plot_grid(ps1Title, SFigS1, ncol = 1, rel_heights = c(0.03, 1))
SFigS1 <- plot_grid(SFigS1, get_legend(ps1a), rel_widths = c(7.5, 1))
SFigS1
ggsave(file.path(outdir, "SFigureSim1.pdf"), width=14, height=16)
```

# Supplementary Figure (unimodal assumption)

This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different sampling distributions of the effect size for non-null tests. The distributions of effect sizes used here were taken from the Stephens (Biostatistics, 2016) paper which originally introduced the ASH method. All effect size distributions are unimodal, with the exception of the appropriately named "bimodal" setting. The effect size distributions from the ASH manuscript have been doubled from their original implementations since TPRs were low (<10%) for all methods at their original magnitude. Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.

```{r SFigureSim2, fig.width = 10, fig.height = 14}
res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist),
                        signal != "bimodal")

## SFigure Sim2 A,B: FDR/TPR of methods across unimodal effect size distribution
ps2a <- genplot(res_ua, cov = alpha, met = "FDR") +
    facet_grid(. ~ signal, labeller = label_parsed)
ps2b <- genplot(res_ua, cov = alpha, met = "TPR") +
    facet_grid(. ~ signal, labeller = label_parsed)

## SFigure Sim2 C,D: FDR/TPR difference of methods
ps2c <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
    facet_grid(. ~ signal, labeller = label_parsed)
ps2d <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
    facet_grid(. ~ signal, labeller = label_parsed)

## create title for SFigure Sim2
ps2Title <- ggdraw() +
    draw_label("Simulation performance across unimodal effect size distributions",
               fontface = 'bold', size = 17)

## pull SFigure Sim2 together
SFigS2 <- plot_grid(ps2a + guides(color = FALSE, linetype = FALSE),
                    ps2b + guides(color = FALSE, linetype = FALSE),
                    ps2c + guides(color = FALSE, linetype = FALSE),
                    ps2d + guides(color = FALSE, linetype = FALSE),
                    labels = LETTERS[1:4], ncol = 1)
SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
SFigS2 <- plot_grid(SFigS2, get_legend(ps2a), rel_widths = c(6.5, 1))
SFigS2
ggsave(file.path(outdir, "SFigureSim2.pdf"), width=10, height=14)
```

# Supplementary Figure (covariate relationship)

This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different function relationships between the informative covariate and null probability (pi0) of a test. In addition to the more deliberately designed sequence of informative covariates studied in the "informativeness" set of simulations described above, we also considered four functional relationships between the covariate and pi0, named "step", "sine", "cosine", and "cubic".

sampling distributions of the effect size for non-null tests. The distributions of effect sizes used here were taken from the Stephens (Biostatistics, 2016) paper which originally introduced the ASH method. All effect size distributions are unimodal, with the exception of the appropriately named "bimodal" setting. The effect size distributions from the ASH manuscript have been doubled from their original implementations since TPRs were low (<10%) for all methods at their original magnitude. Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.

```{r SFigureSim3, fig.width = 12, fig.height = 16}
res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
res_icov <- dplyr::mutate(res_icov, relationship = gsub("informative-", "", setting))

## SFigure Sim3 A: Form of informative covariate and pi0 relationship
pi0form <- tibble(x = seq(0, 1, .01))
pi0form <- dplyr::mutate(pi0form,
                         cosine = pi0_cosine(0.9)(x),
                         cubic = pi0_cubic(0.9)(x),
                         sine = pi0_sine(0.9)(x),
                         step = pi0_step(0.9)(x))
pi0form <- tidyr::gather(pi0form, relationship, pi0, -x)
ps3a <- ggplot(pi0form, aes(x = x, y = pi0)) +
    geom_line() +
    expand_limits(y = c(0.5, 1)) + 
    theme_classic() +
    scale_x_continuous(breaks = seq(0, 1, .2)) + 
    scale_y_continuous(labels = scales::percent) + 
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    facet_grid(. ~ relationship)

## SFigure Sim3 B,C: FDR/TPR of methods across varying relationships
ps3b <- genplot(res_icov, cov = alpha, met = "FDR") +
    facet_grid(. ~ relationship, labeller = label_parsed)
ps3c <- genplot(res_icov, cov = alpha, met = "TPR") +
    facet_grid(. ~ relationship, labeller = label_parsed)

## SFigure Sim3 D,E: FDR/TPR difference of methods
ps3d <- genplot(res_icov, cov = alpha, type = "diff", met = "FDR") +
    facet_grid(. ~ relationship, labeller = label_parsed)
ps3e <- genplot(res_icov, cov = alpha, type = "diff", met = "TPR") +
    facet_grid(. ~ relationship, labeller = label_parsed)

## create title for SFigure Sim3
ps3Title <- ggdraw() +
    draw_label("Simulation performance across informative covariate relationship",
               fontface = 'bold', size = 20)

## pull SFigure Sim3 together
SFigS3 <- plot_grid(ps3a + guides(color = FALSE, linetype = FALSE),
                    ps3b + guides(color = FALSE, linetype = FALSE),
                    ps3c + guides(color = FALSE, linetype = FALSE),
                    ps3d + guides(color = FALSE, linetype = FALSE),
                    ps3e + guides(color = FALSE, linetype = FALSE),
                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
SFigS3 <- plot_grid(ps3Title, SFigS3, ncol = 1, rel_heights = c(0.03, 1))
SFigS3 <- plot_grid(SFigS3, get_legend(ps3b), rel_widths = c(6.5, 1))
SFigS3
ggsave(file.path(outdir, "SFigureSim3.pdf"), width=12, height=16)
```
